Autochecking 2014 NEON phenology data
========================================================
#fixed the ctfiles by domain so D1 is distinct from D1X
#added test for blanks in addDate, for some reason we are getting blank rows in here now
#added test for only one perIndiv record pertagID
#the code now deletes them, but we should fix this as it will cause probs for CI

#still to be done once we get the date format straightened out
#per month, per site, calculate mean, min and max return interval for sampling w.some data entered in any of the columns
#count of # n-y and y-n transitions per phenophase per year per site
#spit out list of species sampled per site.



#per species per site per year, print out min, max mean for all of the annual quantitative meas.

```{r setup so markdown writes nice latex-y looking tables, echo=FALSE}
# code borrowed from here:
#  http://rstudio-pubs-static.s3.amazonaws.com/13339_fdb3c6810e95476fb60c98a657cd00cc.html
require (knitr)
require (xtable)

default_output_handler <- knit_hooks$get("output")

knit_hooks$set(output = function(x, options) {
    if (options["results"] == "jax") {
        sprintf("\\[\n %s\n \\]\n", x)
    } else if ((options["results"] == "hold") && (options["holdjax"] == TRUE)) {
        sprintf("\\[\n %s\n \\]\n", x)
    } else {
        default_output_handler(x, options)
    }
})

#Options templates for cleaner markdown.

opts_template$set(jax = list(comment = "", results = "jax"))
opts_template$set(jaxasis = list(comment = "", results = "asis", echo = FALSE))
opts_template$set(jaxhold = list(comment = "", results = "hold", holdjax = TRUE))

#And a modified toLatex.xtable with two extra settings.

toMathJax.xtable <- function(object, ...) {
    dotArgs <- list(...)
    dotArgs$x <- object
    dotArgs$type <- "latex"
    dotArgs$print.results <- FALSE
    dotArgs$floating <- FALSE
    dotArgs$tabular.environment <- "array"
    z <- do.call("print.xtable", dotArgs)

    z <- strsplit(z, split = "\n")[[1]]
    class(z) <- "Latex"
    z
}

```



```{r make vector of relevant csvs in dropbox,  include=FALSE, cache=FALSE}
require (plyr)
require (stringr)

#set this to the path to your speclist
pathtospeclist<-'N:/common/TOS/ATBDs/Lookups'
#set this to the dropbox
pathtocsvs<-'Z:/'
#set this to the prefix for the sheets in your module
myPrefix<-'phe'


Filelist<-list.files(pathtocsvs) #list all the files
Filelist<-Filelist[grep (myPrefix,Filelist)] #subset to just the ones in your module
Filelist<-Filelist[grep ('.csv',Filelist)] #subset to just the uploaded csvs
```

```{r summary of number of uploaded tables containing text DXX.csv per domain, echo=FALSE}
myDomains<-paste ('D', c(1:20), sep="")
mySummary<-data.frame(Domain=myDomains)
mySummary$num.tables<-rep(NA, nrow(mySummary))
for (i in 1:nrow(mySummary)){
  mySummary$num.tables[i]<-length(grep(paste(mySummary$Domain[i], '.csv', sep=''),Filelist))
  }

```

Summary table of the number of uploaded DXX.csv files per domain in the specified module

```{r print summaries, echo=FALSE, opts.label="jax"}
toMathJax.xtable(xtable(mySummary))

Filesnamedwrong<-list.files(pathtocsvs) #list all the files
#subset to just the ones in your module
Filesnamedwrong<-Filesnamedwrong[grep (myPrefix,Filesnamedwrong, ignore.case=T)] 
Filesnamedwrong<-Filesnamedwrong[-grep ('.csv',Filesnamedwrong)]

toMathJax.xtable(xtable(data.frame(extrafiles<-Filesnamedwrong)))

```

Full list of uploaded files per domain.  If there are not the expected
number of files per domain in the summary above, check out this list to figure
out what is missing.

```{r print filelist, echo=FALSE, opts.label="jax"}
toMathJax.xtable(xtable (data.frame(filesInZDrive=Filelist)), include.rownames=F)

```

```{r phen errorcheck, echo=FALSE}
#grab the master phenology lookup table
speclist<-read.csv (paste (pathtospeclist, '/','L_phe_speciesList.csv',
                           sep=''))

phencheck<-function(i)  {
  #i<-mySummary$Domain[mySummary$num.tables>0][1]
  cat (paste ('\n', '\n', '\n', 'checking Domain', i, '\n', '\n', '\n'))
  cat (paste('perIndiv checks', i, '\n'))
  perIndiv<-NULL
  if (!file.exists(paste (pathtocsvs, '/','phe_perindividual_', i, '.csv', 
                          sep=''))){
    cat (paste('no perIndiv file for', i, '\n', '\n'))
    }  else{
    perIndiv<-read.csv(paste (pathtocsvs, '/','phe_perindividual_', i, '.csv', 
                              sep=''))
    if (!is.null(perIndiv)){
      #check for all blank rows in perIndiv
      #print out the blank rows, and then delete them
      nodates<-which(is.na(perIndiv$addDate))
      if (length(nodates)>0){
        cat ('\n', paste('these rows have no dates and are not subjected to further testing', i,'\n'))
        print (perIndiv[nodates,])
        perIndiv<-perIndiv[-nodates,]
        }
     cat ('\n', paste('range of 90 degree distance', i, '\n'))
      if (min (perIndiv$ninetyDegreeDistance)>1&
            max(perIndiv$ninetyDegreeDistance)<10) {
        print ('looks good')
        } else {
          print (range(perIndiv$ninetyDegreeDistance))
          }
      
      cat ('\n', paste('test for <2 perIndiv record per individual per year', i, '\n'))
      ctrecords<-ddply(perIndiv, c('tagID'), summarize, nrecords=length('tagID'))
      probindivs<-ctrecords$tagID[ctrecords$nrecords!=1]
      if (length(probindivs)>0){
        cat ('these perIndiv records are duplicates or missing')
        print (perIndiv[perIndiv$tagID%in%probindivs,])
      } else {
        print ('looks good')
      }
            
      cat ('\n', paste('number of individuals selected for sampling in site', i,
                       '\n'))
      indivscounted<-ddply (perIndiv, c('plotID'), summarize, 
                            nindiv=length(unique(tagID)))
      if (min(indivscounted$nindiv==90)) {
        print ('looks good')
        } else {
          print (indivscounted, row.names=F)
          }
      cat ('\n', paste('test transect m <800', i, '\n'))
      if (min (perIndiv$transectMeter)>0&max(perIndiv$transectMeter)<800) {
        print ('looks good')
        } else {
          print (range(perIndiv$transectMeter))
          }
      
      cat ('\n', paste('growthform check', i, '\n'))
      
      #this joins the speclist to the perIndiv and makes sure we know 
      #what species each individual is based on it's code
      myGFcheck<-perIndiv   
        
      myGFcheck<-merge(myGFcheck, speclist, all.x=T)
      if (min (!is.na(myGFcheck$scientificName))){
        print ('looks good')
        } else {
          print (unique (myGFcheck[,c('taxonID', 'growthForm',
                                      'scientificName')]))
          }    
      
      cat ('\n', 'remarks perIndiv', i, '\n')
      print (unique(perIndiv$remarks))
      }
    }
  cat (paste('\n', 'statint checks', i, '\n'))
  statint<-NULL
  if (!file.exists(paste (pathtocsvs, '/','phe_statusintensity_', i, '.csv', 
                          sep=''))){
    cat (paste('no statusintensity file for', i, '\n', '\n'))
    }
  else{
    statint<-read.csv(paste (pathtocsvs, '/','phe_statusintensity_', i, '.csv',
                             sep=''))
    if (!is.null(statint)){
      cat ('\n', 'unique status', i, '\n')
      print (unique(c(as.character(statint$p1Status),
                      as.character(statint$p2Status),
                      as.character(statint$p3Status),
                      as.character(statint$p4Status),
                      as.character(statint$p5Status),
                      as.character(statint$p6Status))))
      
      cat ('\n', 'check trouble identifying phenophases', i, '\n')
      recognizephen<-data.frame(status=c('p1Status', 'p2Status', 'p3Status',
                                         'p4Status', 'p5Status', 'p6Status'),
                                propquestionmark=NA)   
      for (j in 1:nrow(recognizephen)){
        dat<-statint[,which(names(statint)==recognizephen$status[j])]
        recognizephen$propquestionmark[j]<-length(dat[dat=='?'])/length(dat[dat%in%c('y', 'n')])
        }
      recognizephen$propquestionmark[!is.finite(recognizephen$propquestionmark)]<-NA
      print (recognizephen)
      
      #make sure everything with a y has an intensity except: P1 intensity 
      #for forbs and grasses does not exist, everything else does, except p6
      
      #first figure out the species for each tagID
      statint<-merge(statint, myGFcheck[,c('tagID', 'growthForm', 
                                           'scientificName')], all.x=T) 
      #P1 intensity for forbs and grasses does not exist
      probs<-NULL
      probs<-c(probs, which(statint$p1Status=='y'&is.na(statint$p1Intensity)&
                              statint$growthForm%in%c('DBL', 'EBL', 'EC', 'DC',
                                                      'Pine')))
      probs<-c(probs,which(statint$p2Status=='y'&is.na(statint$p2Intensity)&
                 statint$growthForm%in%c('DBL', 'EBL', 'EC', 'Pine')))
      probs<-c(probs,which(statint$p3Status=='y'&is.na(statint$p3Intensity)&
                 statint$growthForm%in%c('DBL', 'DC', 'GRS')))
      probs<-c(probs,which(statint$p4Status=='y'&is.na(statint$p4Intensity)))
      probs<-c(probs,which(statint$p5Status=='y'&is.na(statint$p5Intensity)&
                             statint$growthForm%in%c('DBL', 'DC')))
      if (length (probs)==0){
        cat ('\n', 'all statusus have intensities', i, '\n')
        }
      
      if (length (probs)>0){
        cat ('\n', 'these statusus have no intensities', i, '\n')
        print (statint[probs,])
        }
      
      
      #match phenophase to suite of possible phenophases
      cat ('\n', 'check growth form validity', i, '\n')
      probs<-NULL
      probs<-c(probs, which(statint$p1Status%in%c('y', 'n', '?')&
                              statint$growthForm =='Cactus'))
      probs<-c(probs, which(statint$p2Status%in%c('y', 'n', '?')&
                              statint$growthForm %in%c('DC', 'Forb', 'GRS',
                                                       'Cactus')))
      probs<-c(probs, which(statint$p3Status%in%c('y', 'n', '?')&
                              statint$growthForm %in%c('EBL', 'EC', 'Pine')))
      probs<-c(probs, which(statint$p5Status%in%c('y', 'n', '?')&
                              statint$growthForm %in%c('EBL', 'EC', 'Pine',
                                                       'Forb', 'GRS', 'Cactus')))
      probs<-c(probs, which(statint$p6Status%in%c('y', 'n', '?')&
                              statint$growthForm %in%c('EBL', 'EC', 'Pine', 
                                                       'Forb', 'GRS', 'Cactus')))
      if (length (probs)==0){
        cat ('\n', 'All phenophase codes valid for growthForms', i, '\n')
        }
      
      if (length (probs)>0){
        cat ('\n', 'These phenophase codes are invalid for growthForms', i, '\n')
        print (statint[probs,])
        }
      
      
      datebyindiv<-ddply(statint, c('tagID', 'date'), summarize, 
                         nobs=length(tagID))
      if (max(datebyindiv$nobs)==1){
        cat ('\n', 'no duplicate observations per date', i, '\n')
        }
      
      if (length (probs)>0){
        cat ('\n', 'duplicate observations per date', i, '\n')
        myprobs<-merge(statint, datebyindiv[datebyindiv$nobs>0,], all.y=T)
        print (myprobs)
        }
      
      cat ('\n', 'table of intercensus sampling interval in days', i,'\n')
      
      statint<-within(statint, {
        myDates<-paste(substr(date, start=3, stop=4), substr(date,
                    start=5, stop=6), substr(date,
                    start=7, stop=8), sep='/')
        })
      
      statint<-within(statint, myDates<-as.Date(myDates, format = '%y/%m/%d'))
            
      statint$interbout<-NULL
      for (j in 1:nrow(statint)){
        if (length(statint$myDates[statint$tagID==statint$tagID[j]&
                                   statint$myDates<statint$myDates[j]])>0){
          statint$interbout[j]<-statint$myDates[j]-
            max(statint$myDates[statint$tagID==statint$tagID[j]&
                                                statint$myDates<statint$myDates[j]])
        }else {statint$interbout[j]<-NA}
      }
      
      statint$month<-substr(statint$date, start=5, stop=6)      
      print(ddply (statint[!is.na(statint$interbout),], c('plotID', 'month'), summarize, 
                    aveIntercensus=mean(interbout, na.rm=T),
                    minIntercensus= min(interbout, na.rm=T),
                     maxIntercensus= max(interbout, na.rm=T)), row.names=F)  
      
      
      cat ('\n', 'list of species sampled per site', i, '\n')
      print (ddply (statint, c('plotID'), summarize, uniqueSpp=
                      unique(scientificName)), row.names=F)
      
      cat ('\n', 'remarks statint', i, '\n')
      if (length(unique(statint$remarks))>0) {      
      print (unique(statint$remarks))
      }
      else (cat ('no remarks'))
    }
  }
}

```
Phenology check for **Domain 1** 
--------------------------------

```{r phencheck D1, echo=FALSE}
phencheck('D01')

```

Phenology check for **Domain 10** 
--------------------------------

```{r phencheck D0, echo=FALSE}

phencheck('D10')

```


Phenology check for **Domain 11** 
--------------------------------

```{r phencheck D11, echo=FALSE}

phencheck('D11')

```





